---
title: "SCTC Super Bowl 3k/10k"
author: "Aaron Ferrucci"
date: "February 8, 2016"
output: html_document
---

```{r, echo=FALSE,message=FALSE}
source("sctc_sb.R")
```

### 3k results

`r nrow(data3k)` runners participated in the 3k. 

``` {r, echo = FALSE}
ggplot(data3k, aes(x = Gender), xlab = "Gender") +
  geom_bar(stat = "count", aes(fill = Gender)) +
  ggtitle("3k Gender Distribution")
```

The gender distribution for the 3k is tilted somewhat toward female.

```{r, echo=FALSE}
hist3k <- get_age_histogram(data3k) + ggtitle("3k Age Histogram")
print(hist3k)
```

The average age of a 3k runner was `r round(mean(data3k$Age), 1)`. 

The age histogram shows a pretty good age spread, from under 10 to over 80. 

```{r, echo=FALSE}
ggplot(data3k, aes(x = Gender, y = Age, color = Gender)) + geom_boxplot() + ggtitle("3k Age/Gender Boxplot")
```

Plotting gender vs. age as a box plot shows that female runners are generally younger, and there's a bit more variance in age among the male runners.

```{r, echo=FALSE,message=FALSE}
plot3k <- get_plot(data3k, max_age) + ggtitle("3k")
print(plot3k)
```

With a small number of runners, we don't really have meaningful statistics, but nonetheless there seems to be a general upward of run time vs. age. 

### 10k Results

`r nrow(data10k)` runners participated in the 10k.

``` {r, echo = FALSE}
ggplot(data10k, aes(x = Gender), xlab = "Gender") +
  geom_bar(stat = "count", aes(fill = Gender)) +
  ggtitle("10k Gender Distribution")
```

The gender balance for the 10k is tilted more toward male than the 3k was toward female.

```{r, echo=FALSE}
hist10k <- get_age_histogram(data10k) + ggtitle("10k Age Histogram")
print(hist10k)
```

The average age was `r round(mean(data10k$Age), 1)`.

The age histogram for the 10k race is something of a bell curve. There's good representation in the 45-60 group.

```{r, echo=FALSE}
ggplot(data10k, aes(x = Gender, y = Age, color = Gender)) + geom_boxplot() + ggtitle("10k Age/Gender Boxplot")
```

As in the 3k, male 10k runners are older, on average, but it's not as strong a difference.

```{r, echo=FALSE,message=FALSE}
plot10k <- get_plot(data10k, max_age) + ggtitle("10k")
print(plot10k)
```

Again, there's a general upward trend of runtime vs. age, and males are typically faster.

